The human in the loop: volunteer computing as a socio-technical system

### 2016-02-28 Alife16/paper.tex
Revises intro
Please @fergunet @unintendedbear @mariosky revise and, if needed close. Refs #102

\section{Introduction}
\label{introduction}
Nodays there is a big and increasing number of
distributed evolutionary computation software frameworks available on
the market, usually as open source systems \citep{Parejo12Survey}. 
Most of these frameworks have 
two features: they are {\em desktop} applications, that is, they must
be compiled or run directly on the operating system layer and, second,
they assume that the nodes that are going to run the system are under
the control of developers or they at least have access to them.

There is, then, space for frameworks that do not make any of these
assumptions, such as the one presented in this paper, which will be
 named {\sf NodIO}. {\sf NodIO} is basically
% (Paloma) Is the "first" word really needed? 
% (Mario) I guess is more related to fundamentally or primarily but "basically" sounds better
a two tier client/server 
application, in which clients can run on a browser. 
Therefore, they can be embedded in any web-enabled device, including many over which 
developers have no control.

Some time ago, our group faced the problem of diminishing funds for
buying new hardware. This was aggravated by the increasing maintenance costs and
extended downtime resulting from the continuous failures of existing
clusters.
Considering this, we leveraged our experience in the design of web
applications, including JavaScript, since their inception and other
volunteer and
unconventional distributed evolutionary computing systems to 
design and release a new free framework that would allow anyone to
create a volunteer distributed evolutionary computation (EC) experiment using cloud resources as
servers and browsers as clients.

There exist other volunteer computing frameworks written in compiled languages,
such as C++ or Java, that, in principle, can achieve better performance than
JavaScript \citep{2015arXiv151101088M}. However, one of the advantages of our system is that
clients do not need to install anything in their machines, they just
need to open a web page  
in a compatible browser, such as Chrome, Safari or Firefox, which are
available in almost every desktop or mobile device. Other advantage is
that the source code can be directly examined by volunteers, because it is not
compiled as other options such as Java Applets, ActionScript or
ActiveX.
Also, as we will discuss in this paper, it can provide some extra
features such as enhanced security or operating system independence.

When developing volunteer based frameworks, designers must
also consider the whole {\em social} aspect in the design, with
issues related to security, trust and privacy among others. 
The computing system becomes a {\em techno-social system} \citep{vespignani2009predicting}.

In this paper, our intention is to
present the {\sf NodIO} framework and analyze it from two different
angles: the purely technical angle, including the decisions that went
into its features, and the social angle, to account on how the persons
that voluntarily choose to enter the web page running the system
determine its performance and how it can be modeled and, eventually,
predicted. Next we present the state of the art in web-based
volunteer-based systems along with attempts to predict and model its
behavior.

>>>>

\section{Introduction}
\label{introduction}

Some time ago, our group faced the problem of diminishing funds for
buying new hardware. This was aggravated by the increasing maintenance
costs and extended downtime resulting from the continuous failures of
existing clusters.  Considering this, we leveraged our experience in
the design of web applications with JavaScript and other volunteer and
unconventional distributed evolutionary computing systems to design
and release a new free framework that would allow anyone to create a
volunteer distributed evolutionary computation (EC) experiment using
cloud resources as servers and browsers as clients. This framework was
called NodIO \citep{2016arXiv160101607M}. NodIO provides server
infraestructure for volunteer-based distributed evolutionary computing
experiments by providing a chromosome {\em pool}. This pool is used by
clients in browsers and any other using the application programming
interface (API) to put chromosomes and retrieve them, working then as
a loose, asynchronous and {\em ad hoc} connection among all clients
using it.

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate islands that
carry out their own evolutionary algorithms. This is why NodIO is
proposed mainly as a {\em complement} to existent resources such as
desktop systems or laptops. As long as it provides a non-null
computational capability that can help existing resources find the
solution faster it will have found its purpose. The main use case is
someone setting up NodIO in the cloud, writing a fitness evaluation
function and running a client from his or her own computer, but
requesting help in social networks for additional resources. That is
why, in this system,  we have considered the whole {\em social} aspect
in the design, with issues related to security, trust and privacy
among others. The computing system becomes a {\em techno-social
  system} \citep{vespignani2009predicting}. In this paper we are going
to focus on measuring the response of users to experiments, but at the
same time we will also focus on the technical aspects of the server and how
those might change the behavior of users, improving the capability of
the system.

Our research group is committed to open science, and we think this is
a very important part of the techno-social system. By being
transparent, incentives to cheat are reduced and, in fact, we have
detected no issue for the time being. Next we present the state of the art in web-based
volunteer-based systems along with attempts to predict and model its
behavior.
